---
title: "COVID-19 Exploratory Data Analysis"
author: "Iris Jiang"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
   rmdformats::readthedown:
    fig_width: 24
    fig_height: 10
    fig_caption: true
    self_contained: false
    thumbnails: false
    lightbox: false
    gallery: false
    highlight: kate
    toc_depth: 5
    toc_float: true
    css: custom.css
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width = 24)
knitr::opts_chunk$set(fig.height = 10)
knitr::opts_chunk$set(results = "asis")
```


```{r}
## Libraries
library(ggplot2)
library(dplyr)
library(reshape2)
library(data.table)
```


```{r}
# confirmed <- read.csv("time_series_covid19_confirmed_global.csv")
# deaths <- read.csv("time_series_covid19_deaths_global.csv")

# Read data files from website
confirmed <- read.csv("https://data.humdata.org/hxlproxy/api/data-preview.csv?url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_confirmed_global.csv&filename=time_series_covid19_confirmed_global.csv")
deaths <- read.csv("https://data.humdata.org/hxlproxy/api/data-preview.csv?url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_deaths_global.csv&filename=time_series_covid19_deaths_global.csv")


# Reshape data
confirmed <- confirmed %>% 
  rename(Country = Country.Region) %>%
  group_by(Country) %>%
  select(-Lat, -Long) %>%
  ungroup

col_names <- colnames(confirmed)[-(1:2)]
date_list <- substr(col_names, 2, nchar(col_names)-3)
date_list <- gsub("[.]", "/", date_list)
date_list <- paste("2020/", date_list, sep = "")

confirmed_by_country <- rbindlist(lapply(1:length(date_list), function(n){
  confirmed %>% 
    select(Country, !!col_names[n]) %>%
    rename(Cases = !!col_names[n]) %>%
    mutate(Date = as.Date(date_list[n])) %>%
    group_by(Country, Date) %>%
    summarise(Confirmed_Cases = sum(Cases)) %>%
    ungroup
}))  %>%
    group_by(Country) %>%
    mutate(Confirmed_Changes = Confirmed_Cases - lag(Confirmed_Cases)) %>%
    ungroup

deaths <- deaths %>% 
  rename(Country = Country.Region) %>%
  group_by(Country) %>%
  select(-Lat, -Long) %>%
  ungroup

deaths_by_country <- (rbindlist(lapply(1:length(date_list), function(n){
  deaths %>% 
    select(Country, !!col_names[n]) %>%
    rename(Cases = !!col_names[n]) %>%
    mutate(Date = as.Date(date_list[n])) %>%
    group_by(Country, Date) %>%
    summarise(Death_Cases = sum(Cases)) %>%
    ungroup
}))) %>%
    group_by(Country) %>%
    mutate(Death_Changes = Death_Cases - lag(Death_Cases)) %>%
    ungroup

interested_countries <- merge(confirmed_by_country, deaths_by_country, all = T) %>%
  filter(Country %in% c("China", "US", "Canada", "United Kingdom", "Italy", "Germany", "Spain"))

```

## Global Trend

### Confirmed Cases
```{r}
ggplot(interested_countries) +
  geom_line(aes(x = Date, y = Confirmed_Cases, group = Country, color = Country), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Confirmed Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10), labels = scales::scientific)
```

### Confirmed Cases (log scale)
```{r}
ggplot(interested_countries) +
  geom_line(aes(x = Date, y = Confirmed_Cases, group = Country, color = Country), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Confirmed Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(trans='log10', breaks = scales::breaks_log(n = 10, base = 10), labels = scales::scientific)
```

### Daily Change in Confirmed Cases

```{r}
ggplot(interested_countries) +
  geom_line(aes(x = Date, y = Confirmed_Changes, group = Country, color = Country), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Changes of Confirmed Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10), labels = scales::scientific)
```


### Death Cases
```{r}
ggplot(interested_countries) +
  geom_line(aes(x = Date, y = Death_Cases, group = Country, color = Country), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Death Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10), labels = scales::scientific)
```

### Death Cases (log scale)
```{r}
ggplot(interested_countries) +
  geom_line(aes(x = Date, y = Death_Cases, group = Country, color = Country), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Death Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(trans='log10', breaks = scales::breaks_log(n = 10, base = 10), labels = scales::scientific)
```

### Daily Change in Death Cases

```{r}
ggplot(interested_countries) +
  geom_line(aes(x = Date, y = Death_Changes, group = Country, color = Country), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Changes of Death Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

### Daily Change in Death Cases (log scale)

```{r}
ggplot(interested_countries) +
  geom_line(aes(x = Date, y = Death_Changes, group = Country, color = Country), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Changes of Death Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(trans='log10', breaks = scales::breaks_log(n = 10, base = 10), labels = scales::scientific)
```

## Canadian Local Trend

```{r}
confirmed_in_CA <- confirmed %>% 
  filter(Country == "Canada") %>%
  rename(Province = Province.State)

confirmed_in_CA <- rbindlist(lapply(1:length(date_list), function(n){
  confirmed_in_CA %>% 
    select(Province, !!col_names[n]) %>%
    rename(Cases = !!col_names[n]) %>%
    mutate(Date = as.Date(date_list[n])) %>%
    group_by(Province, Date) %>%
    summarise(Confirmed_Cases = sum(Cases)) %>%
    ungroup 
}))  %>%
    group_by(Province) %>%
    mutate(Confirmed_Changes = Confirmed_Cases - lag(Confirmed_Cases)) %>%
    ungroup

death_in_CA <- deaths %>% 
  filter(Country == "Canada") %>%
  rename(Province = Province.State)


death_in_CA <- rbindlist(lapply(1:length(date_list), function(n){
  death_in_CA %>% 
    select(Province, !!col_names[n]) %>%
    rename(Cases = !!col_names[n]) %>%
    mutate(Date = as.Date(date_list[n])) %>%
    group_by(Province, Date) %>%
    summarise(Death_Cases = sum(Cases)) %>%
    ungroup %>%
    mutate( Changes = Death_Cases - lag(Death_Cases)) %>%
    ungroup
})) %>%
    group_by(Province) %>%
    mutate(Death_Changes = Death_Cases - lag(Death_Cases)) %>%
    ungroup

summary_ca <- merge(confirmed_in_CA, death_in_CA, all = T) 

```

### Confirmed Cases
```{r}
ggplot(summary_ca) +
  geom_line(aes(x = Date, y = Confirmed_Cases, group = Province, color = Province), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Confirmed Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

### Confirmed Cases (log scale)
```{r}
ggplot(summary_ca) +
  geom_line(aes(x = Date, y = Confirmed_Cases, group = Province, color = Province), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Confirmed Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(trans='log10', breaks = scales::breaks_log(n = 10, base = 10), labels = scales::scientific)
```

### Daily Change in Confirmed Cases

```{r}
ggplot(summary_ca) +
  geom_line(aes(x = Date, y = Confirmed_Changes, group = Province, color = Province), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Changes of Confirmed Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```


### Death Cases
```{r}
ggplot(summary_ca) +
  geom_line(aes(x = Date, y = Death_Cases, group = Province, color = Province), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Death Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

### Daily Change in Death Cases

```{r}
ggplot(summary_ca) +
  geom_line(aes(x = Date, y = Death_Changes, group = Province, color = Province), size = 1.5) + 
  xlab(NULL) +
  theme(legend.title = element_blank()) + 
  ylab("Changes of Death Cases") +
  scale_x_date(date_breaks = "5 days",  date_labels = "%m-%d", minor_breaks = NULL) + 
  theme(legend.text = element_text(size=16),
        text = element_text(size=20)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```


The data is retrieved from <https://data.humdata.org/dataset/novel-coronavirus-2019-ncov-cases>. The Rmd theme is `readthedown` from the package `rmdformats`.